\documentclass{article}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{enumitem}
\usetikzlibrary{chains,fit,shapes}

\author{Andrew McIsaac}
\title{Introduction to Complexity and Computability: Homework 2}
\date{\today}


\begin{document}
\maketitle

\subsection*{Homework Problem 1}
Show that the following languages are partially decidable:

\begin{enumerate}[label=(\alph*),leftmargin=*]
	\item $ETx = \{\langle M,x \rangle~|~M(x) \downarrow\ \textrm{and at the end
		of computation the tape of $M$ is empty}\}$

		Use a Universal Turing Machine to simulate $M$ on $x$. If $M(x)\downarrow$
		then clear the tape of M and accept.

	\item $ETE = \{\langle M \rangle~|~(\exists~x)(M(x)\downarrow \textrm{and at
		the end of computation the tape of $M$ is empty)}\}$

		Run $ETx$ on every possible input string $x$. Run the first Turing Machine $ETx_{1}$
		on input $x_{1}$ for one step, then the TMs $ETx_{1}$ and $ETx_{2}$ for
		one step on input $x_{1}$ and $x_{2}$ respectively, and so on for every
		single $x$. If there is an $x$ that is accepted by $ETx$, clear the tape
		of M and $ETE$ accepts.
	
\end{enumerate}

\subsection*{Homework Problem 2}
Show that the language $ETx$ is not decidable.

The proof is by contradiction. Assume that $ETx$ is decidable. Then there is a
TM $P$ that decides $ETx$. Then construct a TM $Q$ to decide $L_\mathcal{U}$,
the universal language.

$Q$ proceeds as follows on the input $\langle M,x \rangle$:
\begin{enumerate}
	\itemsep0em
	\item Run $P$ on $\langle M, x \rangle$.
	\item If $P$ rejects, reject.
	\item If P accepts, simulate $M$ on $x$ until it halts and the tape of $M$
		is empty.
	\item If $M$ accepts, accept. If $M$ rejects, reject.
\end{enumerate}

If $P$ decides $ETx$, $Q$ decides $L_\mathcal{U}$. But $L_\mathcal{U}$ is
undecidable, so $ETx$ must also be undecidable.

\end{document}
